Formation and evolution of intermediate-mass ratio inspirals

### 2020-01-16 article.tex

@@

%
% Conclusions: What are the implications of our answer?
% ======================================================
%
We study the formation and evolution of intermediate mass ratio inspirals (IMRIs) triggered by the interactions between two stellar black holes (BHs) and an intermediate black hole (IMBH) inhabiting a globular cluster. We show that the probability for IMRIs formation is an increasing function of the IMBH mass that is well described by a power-law for IMBH masses above $M_\ibh = 10^3\Ms$. We find that IMRIs have a formation probability $\sim 15\%$, regardless of the assumptions made. Our results suggest that in low-density clusters these interactions lead to the formation of stable triples, a fraction of which arrange in a hierarchical configurations and undergo Kozai-Lidov oscillations. In dense clusters, instead, the most frequent configuration is unstable, with the IMBH and the closest BH being strongly affected by the chaotic perturbations exerted by the third BH. 
We calculate IMRIs horizon redshift for several ground- and space-based gravitational waves (GWs) observatories. Combining this information with our results and observational limits on the cosmological distribution of galaxies and star clusters, we infer the IMRI merger rate for different detectors. Assuming a 4 yr observation timespan, we infer a merger rate of 1-4 events yr$^{-1}$ for LIGO in the IMRIs mass range $100-250\Ms$ up to redshift $z=0.1-0.6$, 1-30 yr$^{-1}$ for LISA ($M_{\rm imri} > 10^3\Ms$) up to redshift $z=1.5$. The next generation of detectors will allow us to observe IMRI out to the peak of GC formation and even beyond, with an expected rate of $200-2000$ yr$^{-1}$ for ET ($M_{\rm imri} < 10^3\Ms$) and $250-4400$ yr$^{-1}$ for DECIGO ($M_{\rm imri} < 60 - 5\times 10^4 \Ms$). We show that a considerable number of sources can be observed are multiband thus can be observed with both low- and high-frequency detectors, especially in the case of future detectors. Although highly eccentric at formation, we find that only $\sim 1\%$ of IMRI will preserve a high eccentricity when entering the mHz frequency band. 
\end{abstract}

@@

To explore the dependence between IMRI candidates eccentricity at formation and IMBH mass, we take the eccentricity directly from the last snapshot for models targeted as ``Disrupted'' and ``Unstable'', while for ``Hierarchical'' we record the maximum eccentricity. We then calculate the mean eccentricity value $\langle e_\gw\rangle$  in different mass bins, making the same for the subset of IMRI candidates, as shown in Figure \ref{fig:f4}. In general, it seems that, regardless of the set, IMRI candidates with masses below $5\times10^3\Ms$ have, on average, $\langle e_\gw\rangle$ larger compared to the whole simulations sample, while this trend reverses for heavier IMRIs. 
The average eccentricity increases in the $10^2-5\times10^3\Ms$ mass range, peaking at around $500-1000\Ms$, and than decreases rapidly at increasing the IMBH mass.This implies that the evolution of the inner binary leads the eccentricity to achieve larger values in correspondence of lighter IMBHs. We find that IMBHs with masses typical of star clusters $10^3<M_\ibh/\Ms<10^4$ are expected to form, on average, extremely eccentric IMRIs, while the eccentricity at formation attains more moderate values at IMBH masses typical of low-mass dwarf galaxies and nuclear clusters.
While affecting the eccentricity, the IMBH mass does not have a significant effect on the companion BH mass. Indeed, IMRIs light component masses follows the assumed BH mass distribution.

>>>>

\newcommand{\manuel}{\textcolor{red}}
\newcommand{\pau}{\textcolor{blue}}
\newcommand{\xian}{\textcolor{green}}

@@

%
% Conclusions: What are the implications of our answer?
% ======================================================
%
\manuel{Put your comments by using \textbackslash + YOUR NAME + $\{$TEXT$\}$}\\
We study the formation and evolution of intermediate mass ratio inspirals (IMRIs) triggered by the interactions between two stellar black holes (BHs) and an intermediate black hole (IMBH) inhabiting a globular cluster. We show that the probability for IMRIs formation is an increasing function of the IMBH mass that is well described by a power-law for IMBH masses above $M_\ibh = 10^3\Ms$. We find that IMRIs have a formation probability $\sim 15\%$, regardless of the assumptions made. Our results suggest that in low-density clusters these interactions lead to the formation of stable triples, a fraction of which arrange in a hierarchical configurations and undergo Kozai-Lidov oscillations. In dense clusters, instead, the most frequent configuration is unstable, with the IMBH and the closest BH being strongly affected by the chaotic perturbations exerted by the third BH. 
We calculate IMRIs horizon redshift for several ground- and space-based gravitational waves (GWs) observatories. Combining this information with our results and observational limits on the cosmological distribution of galaxies and star clusters, we infer the IMRI merger rate for different detectors. Assuming a 4 yr observation timespan, we infer a merger rate of 1-4 events yr$^{-1}$ for LIGO in the IMRIs mass range $100-250\Ms$ up to redshift $z=0.1-0.6$, 1-30 yr$^{-1}$ for LISA ($M_{\rm imri} > 10^3\Ms$) up to redshift $z=1.5$. The next generation of detectors will allow us to observe IMRI out to the peak of GC formation and even beyond, with an expected rate of $200-2000$ yr$^{-1}$ for ET ($M_{\rm imri} < 10^3\Ms$) and $250-4400$ yr$^{-1}$ for DECIGO ($M_{\rm imri} < 60 - 5\times 10^4 \Ms$). We show that a considerable number of sources can be observed are multiband thus can be observed with both low- and high-frequency detectors, especially in the case of future detectors. Although highly eccentric at formation, we find that only $\sim 1\%$ of IMRI will preserve a high eccentricity when entering the mHz frequency band. 
\end{abstract}

@@

To explore the dependence between IMRI candidates eccentricity at formation and IMBH mass, we take the eccentricity directly from the last snapshot for models targeted as ``Disrupted'' and ``Unstable'', while for ``Hierarchical'' we record the maximum eccentricity. We then calculate the mean eccentricity value $\langle e_\gw\rangle$  in different mass bins, making the same for the subset of IMRI candidates, as shown in Figure \ref{fig:f5}. In general, it seems that, regardless of the set, IMRI candidates with masses below $5\times10^3\Ms$ have, on average, $\langle e_\gw\rangle$ larger compared to the whole simulations sample, while this trend reverses for heavier IMRIs. 
The average eccentricity increases in the $10^2-5\times10^3\Ms$ mass range, peaking at around $500-1000\Ms$, and than decreases rapidly at increasing the IMBH mass.This implies that the evolution of the inner binary leads the eccentricity to achieve larger values in correspondence of lighter IMBHs. We find that IMBHs with masses typical of star clusters $10^3<M_\ibh/\Ms<10^4$ are expected to form, on average, extremely eccentric IMRIs, while the eccentricity at formation attains more moderate values at IMBH masses typical of low-mass dwarf galaxies and nuclear clusters.
While affecting the eccentricity, the IMBH mass does not have a significant effect on the companion BH mass. Indeed, IMRIs light component masses follows the assumed BH mass distribution.
